### 2020-10-15 latex/acrobot.tex
Edits to describe the phase space explicitly

Defining \(M(q) := D_s(q)\) and \(V(q) := P_s(q)\), we find the conjugate of momenta 
is \(p = (p_u, p_a) = M(q)\dot{q}\).
The dynamics in \((q,p)\) coordinates are given by
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p \\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) \\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
where the inverse inertia matrix is
\begin{equation}\label{eqn:Minv}
    \Minv(q) = \frac{1}{ml^2\left(2-c_a^2\right)}
    \begin{bmatrix}
        1 &
        -\left(1+c_a\right) \\
        -\left(1+c_a\right) &
        3+2c_a
    \end{bmatrix}
    .
\end{equation}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates, allowing us once again to
apply the theory of VNHCs from Chapter \ref{ch:vnhcs}.

Let us define the VNHC \(h(q,p) = q_a - f(q_u,p_u)\) of order 1, where
\(f \in C^2\left(\mathbb{S}^1 \times \R; \R\right)\).
To enforce regularity of \(h(q,p)\), we note that 
\(\nabla_{q_u}\Minv(q) = \Zmat{2\times 2}\).
Examining Theorem \ref{thm:vnhc-regularity}, our VNHC will be regular
if the regularity matrix
\[
    dh_q \Minv(q) \begin{bmatrix}0 \\ 1 \end{bmatrix}
    ,
\]
is of full rank on the constraint manifold \(\Gamma\).
Since 
\[
    dh_q = \begin{bmatrix}
        -\partial_{q_u} f & 1
    \end{bmatrix}
    ,
\]
the regularity matrix evaluates to the scalar equation
\[
    \frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}
\]
which is full rank if and only if the numerator does not
change sign.
A sufficient condition for this is when \(f\) is a function solely of \(p_u\),
so that \(\partial_{q_u}f = 0\).
If this is the case, the regularity matrix is always positive and the VNHC 
\(h(q,p)\) is regular.

The acrobot is a noticeably more complex system than the VLP, as
the dynamics are coupled because \(\Minv(q)\) is not diagonal.
Because of this, the constrained dynamics of an arbitrary VNHC may not be easy
to determine analytically, which will impact our ability to prove energy
injection analytically.
In the rest of this chapter, our goal is to design the function \(f(q_u,p_u)\)
based on the natural human motion of a gymnast, with the caveat that the
resulting VNHC be regular and that we can prove the constrained dynamics will
inject energy into the acrobot.

>>>>

Defining \(M(q) := D_s(q)\) and \(V(q) := P_s(q)\), we find the conjugate of momenta 
is \(p = (p_u, p_a) = M(q)\dot{q}\).
The dynamics in \((q,p)\) coordinates are given by
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p \\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) \\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
where the inverse inertia matrix is
\begin{equation}\label{eqn:Minv}
    \Minv(q) = \frac{1}{ml^2\left(2-c_a^2\right)}
    \begin{bmatrix}
        1 &
        -\left(1+c_a\right) \\
        -\left(1+c_a\right) &
        3+2c_a
    \end{bmatrix}
    .
\end{equation}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\[
    \mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a = 
    \mathbb{S}^ \times \mathbb{S}^1
    ,
\]
and
\[
    \mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
    = \R \times \R
    .
\]
This allowg us once again to
apply the theory of VNHCs from Chapter \ref{ch:vnhcs}.

Let us define the VNHC \(h(q,p) = q_a - f(q_u,p_u)\) of order 1, where
\(f \in C^2\left(\mathcal{Q}_u \times \mathcal{P}_u; \mathcal{Q}_a\right)\).
To enforce regularity of \(h(q,p)\), we note that 
\(\nabla_{q_u}\Minv(q) = \Zmat{2\times 2}\).
Examining Theorem \ref{thm:vnhc-regularity}, our VNHC will be regular
if the regularity matrix
\[
    dh_q \Minv(q) \begin{bmatrix}0 \\ 1 \end{bmatrix}
    ,
\]
is of full rank on the constraint manifold \(\Gamma\).
Since 
\[
    dh_q = \begin{bmatrix}
        -\partial_{q_u} f & 1
    \end{bmatrix}
    ,
\]
the regularity matrix evaluates to the scalar equation
\begin{equation}\label{eqn:regularity-matrix-acrobot}
    \frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}
    ,
\end{equation}
which is full rank if and only if the numerator does not
change sign.

A sufficient condition for regularity is when \(f\) is a function solely of \(p_u\),
because \(\partial_{q_u}f = 0\)
and (\ref{eqn:regularity-matrix-acrobot}) becomes strictly positive for all
values of \(q_a\).

The acrobot is noticeably more complex than the VLP, as
the dynamics of \((q_u,p_u)\) and \((q_a,p_a)\) are coupled through \(\Minv(q)\).
Because of this, the constrained dynamics of an arbitrary VNHC may not be easy
to write out.
In the rest of this chapter, our goal is to design the function \(f(q_u,p_u)\)
based on the natural human motion of a gymnast, with the caveat that the
resulting VNHC be regular and that we can prove the constrained dynamics will
inject energy into the acrobot.
